---
title: "Example of Insulin Curves"
author: "Kevin Cummiskey"
date: "3/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

The following code imports data from the study "Macronutrients and Body Fat Accumulation: A Mechanistic Feeding Study" publicly available at https://osf.io/m6v73/.  

The data files are:

* fb4_use.csv

* fs2_use.csv

See the website for a data dictionary.  The data files contain 37 variables on each subject (demographics, biomarkers, etc.).

### Import the data

```{r}
library(tidyverse)
library(knitr)

#import fb4_use.csv and fs2_use.csv
data1 <- read_csv("https://osf.io/rmpnw/download", col_names = TRUE)
data2 <- read_csv("https://osf.io/qcwft/download", col_names = TRUE)

data <- rbind(data1, data2)

names(data)
```

The subjects were randomized to three test diets:

* Hypocaloric Very-Low-Carbohydrate (VLC)

* High Carbohydrate-Low Sugar (HC-LS)

*  High Carbohydrate-High Sugar (HC-HS)

```{r}
#add subject's treatment group information
subject_info <- read_csv("https://osf.io/3ctnm/download",
                         col_names = TRUE)
subject_info <- subject_info %>% 
  select(randomization, subjectid) %>%
  distinct()
  
data <- data %>% left_join(subject_info,
                           by = c("studyid" = "subjectid"))
#remove subjects not randomized; convert to factor variable
data <- data %>% 
  filter(!is.na(randomization)) %>% 
  mutate(randomization = factor(randomization))
```

### Convert to tidy format for analysis

```{r}
#extract just the insulin measurements and convert to long format
data_BSL <- data %>% 
  select(studyid, randomization, starts_with("BSL")) %>% 
  select(-BSL_HOMA_IR) %>% 
  pivot_longer(cols = -c(studyid, randomization), 
               names_to = "time(min)", 
               values_to = "BSL_insulin")

#extract the time of the measurement (minutes)
#convert N/A (baselines) to zero
data_BSL <- data_BSL %>% 
  mutate(`time(min)` = str_extract(`time(min)`,"\\d+")) %>% 
  mutate(`time(min)` = as.integer(`time(min)`)) %>% 
  mutate(`time(min)` = replace_na(`time(min)`, 0)) 

data_BSL %>% 
  head(10) %>% 
  kable(caption = "First 10 rows of the data")

#save both wide and long formats
#data_BSL %>% 
#  write_csv("./output/BSL_insulin_long.csv")

#data %>% 
#  write_csv("./output/BSL_insulin_wide.csv")

```

### Plot the insulin curves

```{r}
data_BSL %>% 
  ggplot(aes(x = `time(min)`, 
             y = BSL_insulin, 
             group = studyid,
             color = randomization)) +
  labs(y = "OGTT insulin (uIU/mL)",
       color = "treatment group") +
  geom_line()
```


### Future Work

* fit models to each curve, extract parameters

* use parameters in follow-on analyses (for example, are the treatments, covariates, or outcomes related to the parameters.)

